---
title: "coursera5project1"
output: html_document
---

## R Markdown

First, download the data and read it into r
```{r data}

download.file(url = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", destfile = "./data.zip")
unzip(zipfile = "./data.zip")
activity <- read.csv(file = "activity.csv")

```

## Number of Steps each day

Histogram:

```{r histOfSteps }
stepsPerDay <- xtabs(activity$steps~activity$date)
stepsPerDayData <- data.frame(stepsPerDay)
hist(stepsPerDayData$Freq, xlab = "Number of Steps", ylab = "Number of Days", main = "Histogram of Steps Each Day")

```

Summary Statistics:

```{r summarySteps}

summary(stepsPerDayData$Freq)

```


## Time series of Average Steps:

```{r timeSeries }

library(plyr)
aveStpsPerInt <- ddply(activity, .(interval), summarize, mean = mean(steps, na.rm = TRUE))
plot(x = aveStpsPerInt$interval, y = aveStpsPerInt$mean, xlab = "Interval", ylab = "Average Steps")

```

##The five-minute Interval that Contains the Highest Average Steps

```{r max interval}

aveStpsPerInt[which.max(aveStpsPerInt$mean),]

```

##Inputting Missing Data:
1.  Find NA's in original Data
2.  Set the missing NA data to the average steps taken for the cooresponding interval

```{r missing data}
naInputter = function(x,y) {
  
  ##find na's in activity
  for(i in 1:nrow(x)){
    if(is.na(x[i,1])){
      ##match na to ave mean for cooresponding interval
      x[i,1] <- y[which(y$interval == x[i,3]),2]
  
    }
  }
  x
}
activity2 <- naInputter(activity, aveStpsPerInt)

```

##Histogram of Total Number of Steps Taken Each Day After Missing Values are Inputted
  
```{r hist}
stepsPerDay2 <- xtabs(activity2$steps~activity2$date)
hist(stepsPerDay2, xlab = "Number of Steps", ylab = "Number of Days", main = "Histogram of Steps with Missing Data")
```

Panel Plot Comparing the Average Number of Steps Taken per 5-Minute Interval Across Weekdays and Weekends

```{r Panel Plot}
library(dplyr)
##find weekdays within activity data
weekendSep <- function(x){
  x$weekday <- weekdays(as.Date(x$date))
  x <- mutate(x,dayInd = "null")
  weekends <- c("Saturday","Sunday")
  for(i in 1:nrow(x)){
    if(x$weekday[i] == weekends[1] || x$weekday[i] == weekends[2]){
      
      x$dayInd[i] = "Weekend"
    }
    else{
      x$dayInd[i] = "Weekday"
    }
  }
  x
}
activity2 <- weekendSep(activity2)

weekdayData <- filter(activity2, dayInd == "Weekday")
weekendData <- filter(activity2, dayInd == "Weekend")
weekdayData <- group_by(weekdayData,interval)
weekendData <- group_by(weekendData,interval)
aveWeekday <- ddply(weekdayData, .(interval), summarize, mean = mean(steps, na.rm = TRUE))
aveWeekend <- ddply(weekendData, .(interval), summarize, mean = mean(steps, na.rm = TRUE))
par(mfrow = c(2,1), mar = c(2,2,2,2))
plot(aveWeekday, main = "Weekdays")
plot(aveWeekend, main = "Weekends")

```






